---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

```{r include=FALSE, echo=FALSE}
data<-readRDS("clean_data_0423.rds")
dim(data)  # 2110, 24
unique(is.na(data))
colnames(data)
library(survival)
library(survminer)
library(KMsurv)
```

# categorical BOOZE_q

```{r echo=FALSE, fig.width=8}
km_fit_1<-survfit(Surv(AGEYRS, AGEDIE, cancer_death)~BOOZE_q, data=data)
ggsurvplot(km_fit_1, data=data, risk.table=FALSE)
```

\newpage

## ~ BOOZE_q
```{r echo=FALSE, fig.height=4}
cox_fit_1<-coxph(Surv(AGEYRS, AGEDIE, cancer_death)~BOOZE_q, data=data)
summary(cox_fit_1)
resi1<-cox.zph(cox_fit_1)
resi1
plot(resi1, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_1)[1])
```

\newpage

## ~ BOOZE_q + SEX
```{r echo=FALSE, fig.height=4}
cox_fit_2<-coxph(Surv(AGEYRS, AGEDIE, cancer_death)~BOOZE_q+SEX, data=data)
summary(cox_fit_2)
resi2<-cox.zph(cox_fit_2)
resi2
plot(resi2, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_2)[1])
```

\newpage

## ~ BOOZE_q + SEX + BOOZE_q * SEX
```{r echo=FALSE, fig.height=4}
cox_fit_3<-coxph(Surv(AGEYRS, AGEDIE, cancer_death)~BOOZE_q+SEX+BOOZE_q * SEX, data=data)
summary(cox_fit_3)
resi3<-cox.zph(cox_fit_3)
resi3
plot(resi3, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_3)[1])
```

\newpage

## ~ BOOZE_q, stratify by SEX 
```{r echo=FALSE, fig.height=4}
cox_fit_4<-coxph(Surv(AGEYRS, AGEDIE, cancer_death)~BOOZE_q+strata(SEX), data=data)
summary(cox_fit_4)
resi4<-cox.zph(cox_fit_4)
resi4
plot(resi4, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_4)[1])
```

\newpage

## ~ BOOZE_q + SEX + RACE + GRADES + MARRY + SIZE + AVGSMK + SMSA + URBAN
```{r echo=FALSE, fig.height=4}
cox_fit_5<-coxph(Surv(AGEYRS, AGEDIE, cancer_death)~BOOZE_q+SEX+ RACE + GRADES + MARRY + SIZE + AVGSMK + SMSA + URBAN, data=data)
summary(cox_fit_5)
resi5<-cox.zph(cox_fit_5)
resi5
plot(resi5, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_5)[1])
```

\newpage

## ~ BOOZE_q + SEX + BOOZE_q * SEX + RACE + GRADES + MARRY + SIZE + AVGSMK + SMSA + URBAN
```{r echo=FALSE, fig.height=4}
cox_fit_6<-coxph(Surv(AGEYRS, AGEDIE, cancer_death)~BOOZE_q+SEX+BOOZE_q*SEX+ RACE + GRADES + MARRY + SIZE + AVGSMK + SMSA + URBAN, data=data)
summary(cox_fit_6)
resi6<-cox.zph(cox_fit_6)
resi6
plot(resi6, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_6)[1])

coef<-coef(cox_fit_6)[c(3,14)] %>% as.matrix(ncol=1)
vcov<-vcov(cox_fit_6)[c(3,14),c(3,14)]
var=t(coef)%*%vcov%*%coef

beta <- coef(cox_fit_6)[c(3,14)]
SE <- c(coef(summary(cox_fit_6))[3,3], sqrt(var))
lcl <- beta-qnorm(0.975)*SE 
ucl <- beta+qnorm(0.975)*SE
cbind(exp(beta), exp(lcl), exp(ucl))
```

\newpage

## ~ BOOZE_q +BOOZE_q*stratify(SEX) + RACE + GRADES + MARRY + SIZE + AVGSMK + SMSA + URBAN, stratify by SEX 
```{r echo=FALSE, fig.height=4}
cox_fit_7<-coxph(Surv(AGEYRS, AGEDIE, cancer_death)~BOOZE_q + strata(SEX)+ BOOZE_q*strata(SEX) + RACE + GRADES + MARRY + SIZE + AVGSMK + SMSA + URBAN, data=data)
summary(cox_fit_7)
resi7<-cox.zph(cox_fit_7)
resi7
plot(resi7, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_7)[1])

coef<-coef(cox_fit_7)[c(3,13)] %>% as.matrix(ncol=1)
vcov<-vcov(cox_fit_7)[c(3,13),c(3,13)]
var=t(coef)%*%vcov%*%coef

beta <- coef(cox_fit_7)[c(3,13)]
SE <- c(coef(summary(cox_fit_7))[3,3], sqrt(var))
lcl <- beta-qnorm(0.975)*SE 
ucl <- beta+qnorm(0.975)*SE
cbind(exp(beta), exp(lcl), exp(ucl))
```


\newpage

# continuous BOOZE

## ~ BOOZE
```{r echo=FALSE, fig.height=4}
cox_fit_c1<-coxph(Surv(AGEYRS, AGEDIE, cancer_death)~BOOZE, data=data)
summary(cox_fit_c1)
resic1<-cox.zph(cox_fit_c1)
resic1
plot(resic1, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_c1)[1])
```

\newpage

## ~ BOOZE + SEX
```{r echo=FALSE, fig.height=4}
cox_fit_c2<-coxph(Surv(AGEYRS, AGEDIE, cancer_death)~BOOZE+SEX, data=data)
summary(cox_fit_c2)
resic2<-cox.zph(cox_fit_c2)
resic2
plot(resic2, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_c2)[1])
```

\newpage

## ~ BOOZE + SEX + BOOZE * SEX
```{r echo=FALSE, fig.height=4}
cox_fit_c3<-coxph(Surv(AGEYRS, AGEDIE, cancer_death)~BOOZE+SEX+BOOZE * SEX, data=data)
summary(cox_fit_c3)
resic3<-cox.zph(cox_fit_c3)
resic3
plot(resic3, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_c3)[1])
```

\newpage

## ~ BOOZE +BOOZE*SEX, stratify by SEX 
```{r echo=FALSE, fig.height=4}
cox_fit_c4<-coxph(Surv(AGEYRS, AGEDIE, cancer_death)~BOOZE+strata(SEX)+BOOZE*SEX, data=data)
summary(cox_fit_c4)
resic4<-cox.zph(cox_fit_c4)
resic4
plot(resic4, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_c4)[1])
```

\newpage

## ~ BOOZE + SEX + RACE + GRADES + MARRY + SIZE + AVGSMK + SMSA + URBAN
```{r echo=FALSE, fig.height=4}
cox_fit_c5<-coxph(Surv(AGEYRS, AGEDIE, cancer_death)~BOOZE+SEX+ RACE + GRADES + MARRY + SIZE + AVGSMK + SMSA + URBAN, data=data)
summary(cox_fit_c5)
resic5<-cox.zph(cox_fit_c5)
resic5
plot(resic5, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_c5)[1])
```

\newpage

## ~ BOOZE + SEX + BOOZE_q * SEX + RACE + GRADES + MARRY + SIZE + AVGSMK + SMSA + URBAN
```{r echo=FALSE, fig.height=4}
cox_fit_c6<-coxph(Surv(AGEYRS, AGEDIE, cancer_death)~BOOZE+SEX+BOOZE*SEX+ RACE + GRADES + MARRY + SIZE + AVGSMK + SMSA + URBAN, data=data)
summary(cox_fit_c6)
resic6<-cox.zph(cox_fit_c6)
resic6
plot(resic6, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_c6)[1])

coef<-coef(cox_fit_c6)[c(1,10)] %>% as.matrix(ncol=1)
vcov<-vcov(cox_fit_c6)[c(1,10),c(1,10)]
var=t(coef)%*%vcov%*%coef

beta <- coef(cox_fit_c6)[c(1,10)]
SE <- c(coef(summary(cox_fit_c6))[1,3], sqrt(var))
lcl <- beta-qnorm(0.975)*SE 
ucl <- beta+qnorm(0.975)*SE
cbind(exp(beta), exp(lcl), exp(ucl))
```

\newpage

## ~ BOOZE +BOOZE*strata(SEX)+ RACE + GRADES + MARRY + SIZE + AVGSMK + SMSA + URBAN, stratify by SEX 
```{r echo=FALSE, fig.height=4}
cox_fit_c7<-coxph(Surv(AGEYRS, AGEDIE, cancer_death)~BOOZE +BOOZE*strata(SEX)+ strata(SEX) + RACE + GRADES + MARRY + SIZE + AVGSMK + SMSA + URBAN, data=data)
summary(cox_fit_c7)
resic7<-cox.zph(cox_fit_c7)
resic7
plot(resic7, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_c7)[1])

coef<-coef(cox_fit_c7)[c(1,9)] %>% as.matrix(ncol=1)
vcov<-vcov(cox_fit_c7)[c(1,9),c(1,9)]
var=t(coef)%*%vcov%*%coef

beta <- coef(cox_fit_c7)[c(1,9)]
SE <- c(coef(summary(cox_fit_c7))[1,3], sqrt(var))
lcl <- beta-qnorm(0.975)*SE 
ucl <- beta+qnorm(0.975)*SE
cbind(exp(beta), exp(lcl), exp(ucl))
```

# ordinal BOOZE_q_ordinal

## ~ BOOZE_q_ordinal
```{r echo=FALSE, fig.height=4}
cox_fit_o1<-coxph(Surv(AGEYRS, AGEDIE, cancer_death)~as.numeric(BOOZE_q), data=data)
summary(cox_fit_o1)
resio1<-cox.zph(cox_fit_o1)
resio1
plot(resio1, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_o1)[1])
```

\newpage

## ~ BOOZE_q_ordinal + SEX
```{r echo=FALSE, fig.height=4}
cox_fit_o2<-coxph(Surv(AGEYRS, AGEDIE, cancer_death)~as.numeric(BOOZE_q)+SEX, data=data)
summary(cox_fit_o2)
resio2<-cox.zph(cox_fit_o2)
resio2
plot(resio2, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_o2)[1])
```

\newpage

## ~ BOOZE_q_ordinal + SEX + BOOZE_q_ordinal * SEX
```{r echo=FALSE, fig.height=4}
cox_fit_o3<-coxph(Surv(AGEYRS, AGEDIE, cancer_death)~as.numeric(BOOZE_q)+SEX+as.numeric(BOOZE_q) * SEX, data=data)
summary(cox_fit_o3)
resio3<-cox.zph(cox_fit_o3)
resio3
plot(resio3, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_o3)[1])
```

\newpage

## ~ BOOZE_q_ordinal, stratify by SEX 
```{r echo=FALSE, fig.height=4}
cox_fit_o4<-coxph(Surv(AGEYRS, AGEDIE, cancer_death)~as.numeric(BOOZE_q)+strata(SEX), data=data)
summary(cox_fit_o4)
resio4<-cox.zph(cox_fit_o4)
resio4
plot(resio4, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_o4)[1])
```

\newpage

## ~ BOOZE_q_ordinal + SEX + RACE + GRADES + MARRY + SIZE + AVGSMK + SMSA + URBAN
```{r echo=FALSE, fig.height=4}
cox_fit_o5<-coxph(Surv(AGEYRS, AGEDIE, cancer_death)~as.numeric(BOOZE_q)+SEX+ RACE + GRADES + MARRY + SIZE + AVGSMK + SMSA + URBAN, data=data)
summary(cox_fit_o5)
resio5<-cox.zph(cox_fit_o5)
resio5
plot(resio5, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_o5)[1])
```

\newpage

## ~ BOOZE_q_ordinal + SEX + BOOZE_q_ordinal * SEX + RACE + GRADES + MARRY + SIZE + AVGSMK + SMSA + URBAN
```{r echo=FALSE, fig.height=4}
cox_fit_o6<-coxph(Surv(AGEYRS, AGEDIE, cancer_death)~as.numeric(BOOZE_q)+SEX+as.numeric(BOOZE_q)*SEX+ RACE + GRADES + MARRY + SIZE + AVGSMK + SMSA + URBAN, data=data)
summary(cox_fit_o6)
resio6<-cox.zph(cox_fit_o6)
resio6
plot(resio6, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_o6)[1])
```

\newpage

## ~ BOOZE_q_ordinal + RACE + GRADES + MARRY + SIZE + AVGSMK + SMSA + URBAN, stratify by SEX 
```{r echo=FALSE, fig.height=4}
cox_fit_o7<-coxph(Surv(AGEYRS, AGEDIE, cancer_death)~as.numeric(BOOZE_q) + strata(SEX) + RACE + GRADES + MARRY + SIZE + AVGSMK + SMSA + URBAN, data=data)
summary(cox_fit_o7)
resio7<-cox.zph(cox_fit_o7)
resio7
plot(resio7, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_o7)[1])
```




