---
title: "Cox model"
author: "Fuyu Guo; Jueyi Liu; Xinyi qian"
date: "4/27/2022"
output:
  pdf_document: default
  html_document: default
---

This RMD file is corresponding to the Cox models in the main analysis following the second conceputal framework (the study is simplied as a cross-sectional study and time scale is time since entry)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load data and packages
```{r}
library(tidyverse)
library(survival)
dta <- readRDS("clean_data_0425.rds")


dta$fu_time <- ifelse(!is.na(dta$DIE_YR),  dta$DIE_YR - dta$EXAM_YR,
                      dta$LAST_YR-dta$EXAM_YR)
# exclude people whose follow-up time is 0
dta <- dta %>% filter(fu_time > 0)
```

# Main analysis
## Model 1: Crude analysis
```{r}
dta$BOOZE_ord <- dta$BOOZE_q %>% as.numeric()
# categorical version
fit1_q <- coxph(Surv(fu_time, cancer_death) ~ BOOZE_q, data = dta)
fit1_q %>% summary()

# continuous version
fit1_con <- coxph(Surv(fu_time, cancer_death) ~ BOOZE, data = dta)
fit1_con %>% summary()

# ordinal version for p-value for trend
fit1_ord <- coxph(Surv(fu_time, cancer_death) ~ BOOZE_ord, data = dta)
fit1_ord %>% summary()
```



## Model 2: Adjusting for age
```{r}
# categorical version
fit2_q <- coxph(Surv(fu_time, cancer_death) ~ BOOZE_q + AGEYRS, data = dta)
fit2_q %>% summary()

# continuous version
fit2_con <- coxph(Surv(fu_time, cancer_death) ~ BOOZE + AGEYRS, data = dta)
fit2_con %>% summary()

# ordinal version for p-value for trend
fit2_ord <- coxph(Surv(fu_time, cancer_death) ~ BOOZE_ord + AGEYRS, data = dta)
fit2_ord %>% summary()
```


## Model 3: MV-adjusted analysis
```{r}
# categorical version
fit3_q <- coxph(Surv(fu_time, cancer_death) ~ BOOZE_q + AGEYRS +
                      SEX + as.factor(RACE) + GRADES + as.factor(SIZE) + as.factor(MARRY) + AVGSMK + as.factor(SMSA) + URBAN + 
                      HTN_REP + RBC + DRMI + DIAB, data = dta)
fit3_q %>% summary()

# continuous version
fit3_con <- coxph(Surv(fu_time, cancer_death) ~ BOOZE + AGEYRS +
                      SEX + as.factor(RACE) + GRADES + as.factor(SIZE)+ as.factor(MARRY) + AVGSMK + as.factor(SMSA) + URBAN + 
                      HTN_REP + RBC + DRMI + DIAB, data = dta)
fit3_con %>% summary()

# ordinal version for p-value for trend
fit3_ord <- coxph(Surv(fu_time, cancer_death) ~ BOOZE_ord + AGEYRS +
                      SEX + as.factor(RACE) + GRADES + as.factor(SIZE) + as.factor(MARRY) + AVGSMK + as.factor(SMSA) + URBAN + 
                      HTN_REP + RBC + DRMI + DIAB, data = dta)
fit3_ord %>% summary()
```

# Exploring effect modification by sex

## Sex interaction terms in the Cox model
```{r}
dta$SEX <- if_else(dta$SEX == 1, "Male", "Female")

# categorical version
fit4_q <- coxph(Surv(fu_time, cancer_death) ~ BOOZE_q + BOOZE_q:SEX + AGEYRS +
                      SEX + as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN + 
                      HTN_REP + RBC + DRMI + DIAB, data = dta)
fit4_q %>% summary()
coef <- coef(fit4_q)[c(3, 31)] %>% as.matrix(ncol = 1)
vcov <- vcov(fit4_q)[c(3, 31), c(3, 31)]
var <- t(c(1, 1)) %*% vcov %*% c(1, 1)
sd <- sqrt(var)
mean_male <- coef[2,] + coef[1,]
exp(mean_male)
c(mean_male - 1.96*sd, mean_male+ 1.96*sd) %>% exp()
```


```{r}
# Continuous version
fit4_con <- coxph(Surv(fu_time, cancer_death) ~ BOOZE + BOOZE:SEX  + AGEYRS +
                      SEX + as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN + 
                      HTN_REP + RBC + DRMI + DIAB, data = dta)
fit4_con %>% summary()
coef <- coef(fit4_con)[c(1, 27)] %>% as.matrix(ncol = 1)
vcov <- vcov(fit4_con)[c(1, 27), c(1, 27)]
var <- t(c(1, 1)) %*% vcov %*% c(1, 1)
sd <- sqrt(var)
mean_male <- coef[2,] + coef[1,]
exp(mean_male)
c(mean_male - 1.96*sd, mean_male+ 1.96*sd) %>% exp()
```

## Stratified Cox model by sex, and adding an interaction term by SEX and alcohol consumption
```{r}
# Categorical version
fit5_q <- coxph(Surv(fu_time, cancer_death) ~ BOOZE_q + strata(SEX) + BOOZE_q:strata(SEX) + AGEYRS  + as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN + 
                      HTN_REP + RBC + DRMI + DIAB, data = dta)
fit5_q %>% summary()
coef <- coef(fit5_q)[c(3, 30)] %>% as.matrix(ncol = 1)
vcov <- vcov(fit5_q)[c(3, 30), c(3, 30)]
var <- t(c(1, 1)) %*% vcov %*% c(1, 1)
sd <- sqrt(var)
mean_male <- coef[2,] + coef[1,]
exp(mean_male)
c(mean_male - 1.96*sd, mean_male+ 1.96*sd) %>% exp()
```
```{r}
# Continuous version
fit5_cont <- coxph(Surv(fu_time, cancer_death) ~ BOOZE + strata(SEX) + BOOZE:strata(SEX) + AGEYRS  + as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN + 
                      HTN_REP + RBC + DRMI + DIAB, data = dta)
fit5_cont %>% summary()
coef <- coef(fit5_cont)[c(1, 26)] %>% as.matrix(ncol = 1)
vcov <- vcov(fit5_cont)[c(1, 26), c(1, 26)]
var <- t(c(1, 1)) %*% vcov %*% c(1, 1)
sd <- sqrt(var)
mean_male <- coef[2,] + coef[1,]
exp(mean_male)
c(mean_male - 1.96*sd, mean_male+ 1.96*sd) %>% exp()
```


## Subgroup analysis

```{r}
# categorical version
fit6_q_m <- coxph(Surv(fu_time, cancer_death) ~ BOOZE_q + AGEYRS  + as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN + 
                      HTN_REP + RBC + DRMI + DIAB, data = dta[dta$SEX == "Male",])
fit6_q_m %>% summary()

fit6_q_f <- coxph(Surv(fu_time, cancer_death) ~ BOOZE_q + AGEYRS  + as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN + 
                      HTN_REP + RBC + DRMI + DIAB, data = dta[dta$SEX == "Female",])
fit6_q_f %>% summary()


z <- (coef(fit6_q_f)[3] - coef(fit6_q_m)[3])/sqrt(vcov(fit6_q_f)[3,3] + vcov(fit6_q_m)[3,3])
1 - pnorm(abs(z))
```



```{r}
# continuous version
fit6_cont_m <- coxph(Surv(fu_time, cancer_death) ~ BOOZE + AGEYRS  + as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN + 
                      HTN_REP + RBC + DRMI + DIAB, data = dta[dta$SEX == "Male",])
fit6_cont_m %>% summary()

fit6_cont_f <- coxph(Surv(fu_time, cancer_death) ~ BOOZE + AGEYRS  + as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN + 
                      HTN_REP + RBC + DRMI + DIAB, data = dta[dta$SEX == "Female",])
fit6_cont_f %>% summary()


z <- (coef(fit6_cont_f)[1] - coef(fit6_cont_m)[1])/sqrt(vcov(fit6_cont_f)[1,1] + vcov(fit6_cont_m)[1,1])
1 - pnorm(z)
```


# Checking PH-assumption
```{r}
# categorical version
fit3_q <- coxph(Surv(fu_time, cancer_death) ~ BOOZE_q + AGEYRS +
                      SEX + as.factor(RACE) + GRADES + as.factor(SIZE) + as.factor(MARRY) + AVGSMK + as.factor(SMSA) + URBAN + 
                      HTN_REP + RBC + DRMI + DIAB, data = dta)
cox.zph(fit3_q)

```

```{r}
# continuous version
fit3_con <- coxph(Surv(fu_time, cancer_death) ~ BOOZE + AGEYRS +
                      SEX + as.factor(RACE) + GRADES + as.factor(SIZE) + as.factor(MARRY) + AVGSMK + as.factor(SMSA) + URBAN + 
                      HTN_REP + RBC + DRMI + DIAB, data = dta)
cox.zph(fit3_con)
```

# Exploring Non-linearity
```{r}
library(splines)
fit_nl <- coxph(Surv(fu_time, cancer_death) ~ ns(BOOZE, knots = 3) + AGEYRS +
                      SEX + as.factor(RACE) + GRADES + as.factor(SIZE) + as.factor(MARRY) + AVGSMK + as.factor(SMSA) + URBAN + 
                      HTN_REP + RBC + DRMI + DIAB, data = dta)
x <- seq(0,21, 0.5)
loghr <- ns(x, knots = 3) %*% coef(fit_nl)[c(1,2)]
vcov <- vcov(fit_nl)[c(1,2), c(1,2)]
var <- diag(ns(x, knots = 3) %*% vcov(fit_nl)[c(1,2), c(1,2)] %*% t(ns(x, knots = 3)))
se <- sqrt(var)
df <- tibble(x = x,
             y = loghr, 
             ymin = loghr - 1.96*se,
             ymax = loghr + 1.96*se)
ggplot(df) + 
    geom_line(aes(x = x, y = y), color = "red", linetype = "dashed")+
    geom_ribbon(aes(x = x, y = y, ymin = ymin, ymax = ymax),
                alpha = 0.5, fill = "#DB7093") + 
    theme_bw() + 
    labs(x = "Alcohol consumption (drinks/week)",
         y = "log(HR)")
```
