---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

```{r include=FALSE, echo=FALSE}
data<-readRDS("clean_data_0425.rds")
dim(data)  # 9208   29
unique(is.na(data))
colnames(data)

data$fu_time <- ifelse(!is.na(data$DIE_YR),  data$DIE_YR - data$EXAM_YR,
                      data$LAST_YR-dta$EXAM_YR)
# exclude people whose follow-up time is 0
data <- data %>% filter(fu_time > 0)

library(survival)
library(survminer)
library(KMsurv)
```

# categorical BOOZE_q
## Crude analysis
```{r echo=FALSE, fig.width=8}
km_fit_1<-survfit(Surv(AGEYRS, AGE_LAST, cancer_death)~BOOZE_q, data=data)
ggsurvplot(km_fit_1, data=data, risk.table=FALSE)
```

\newpage

## ~ BOOZE_q
```{r echo=FALSE, fig.height=4}
cox_fit_1<-coxph(Surv(AGEYRS, AGE_LAST, cancer_death)~BOOZE_q, data=data)
summary(cox_fit_1)
resi1<-cox.zph(cox_fit_1)
resi1
plot(resi1, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_1)[1])
```


\newpage

## ~ BOOZE_q + SEX + as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN
```{r echo=FALSE, fig.height=4}
cox_fit_5<-coxph(Surv(AGEYRS, AGE_LAST, cancer_death)~BOOZE_q+SEX+ as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN, data=data)
summary(cox_fit_5)
resi5<-cox.zph(cox_fit_5)
resi5
plot(resi5, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_5)[1])
```

\newpage

## ~ BOOZE_q + SEX + BOOZE_q * SEX + as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN
```{r echo=FALSE, fig.height=4}
cox_fit_6<-coxph(Surv(AGEYRS, AGE_LAST, cancer_death)~BOOZE_q+SEX+BOOZE_q*SEX+ as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN, data=data)
summary(cox_fit_6)
resi6<-cox.zph(cox_fit_6)
resi6
plot(resi6, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_6)[1])
# 2 vs 0
coef<-coef(cox_fit_6)[c(3,26)] %>% as.matrix(ncol=1)
vcov<-vcov(cox_fit_6)[c(3,26),c(3,26)]
var=t(c(1,1))%*%vcov%*%c(1,1)

beta <- c(coef(cox_fit_6)[3], sum(coef(cox_fit_6)[c(3,26)]))
SE <- c(coef(summary(cox_fit_6))[3,3], sqrt(var))
lcl <- beta-qnorm(0.975)*SE 
ucl <- beta+qnorm(0.975)*SE
out<-cbind(exp(beta), exp(lcl), exp(ucl))
rownames(out)<-c("M","F")
out
# 0.5-2 vs 0
coef<-coef(cox_fit_6)[c(2,25)] %>% as.matrix(ncol=1)
vcov<-vcov(cox_fit_6)[c(2,25),c(2,25)]
var=t(c(1,1))%*%vcov%*%c(1,1)

beta <- c(coef(cox_fit_6)[2], sum(coef(cox_fit_6)[c(2,25)]))
SE <- c(coef(summary(cox_fit_6))[2,3], sqrt(var))
lcl <- beta-qnorm(0.975)*SE 
ucl <- beta+qnorm(0.975)*SE
out<-cbind(exp(beta), exp(lcl), exp(ucl))
rownames(out)<-c("M","F")
out
# 0-0.5 vs 0
coef<-coef(cox_fit_6)[c(1,24)] %>% as.matrix(ncol=1)
vcov<-vcov(cox_fit_6)[c(1,24),c(1,24)]
var=t(c(1,1))%*%vcov%*%c(1,1)

beta <- c(coef(cox_fit_6)[1], sum(coef(cox_fit_6)[c(1,24)]))
SE <- c(coef(summary(cox_fit_6))[1,3], sqrt(var))
lcl <- beta-qnorm(0.975)*SE 
ucl <- beta+qnorm(0.975)*SE
out<-cbind(exp(beta), exp(lcl), exp(ucl))
rownames(out)<-c("M","F")
out
```

\newpage

## ~ BOOZE_q +BOOZE_q*stratify(SEX) + as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN, stratify by SEX 
```{r echo=FALSE, fig.height=4}
cox_fit_7<-coxph(Surv(AGEYRS, AGE_LAST, cancer_death)~BOOZE_q + strata(SEX)+ BOOZE_q*strata(SEX) + as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN, data=data)
summary(cox_fit_7)
resi7<-cox.zph(cox_fit_7)
resi7
plot(resi7, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_7)[1])
# 2 vs 0
coef<-coef(cox_fit_7)[c(3,25)] %>% as.matrix(ncol=1)
vcov<-vcov(cox_fit_7)[c(3,25),c(3,25)]
var=t(c(1,1))%*%vcov%*%c(1,1)

beta <- c(coef(cox_fit_7)[3], sum(coef(cox_fit_7)[c(3,25)]))
SE <- c(coef(summary(cox_fit_7))[3,3], sqrt(var))
lcl <- beta-qnorm(0.975)*SE 
ucl <- beta+qnorm(0.975)*SE
out<-cbind(exp(beta), exp(lcl), exp(ucl))
rownames(out)<-c("M","F")
out
# 0.5-2 vs 0
coef<-coef(cox_fit_7)[c(2,24)] %>% as.matrix(ncol=1)
vcov<-vcov(cox_fit_7)[c(2,24),c(2,24)]
var=t(c(1,1))%*%vcov%*%c(1,1)

beta <- c(coef(cox_fit_7)[2], sum(coef(cox_fit_7)[c(2,24)]))
SE <- c(coef(summary(cox_fit_7))[2,3], sqrt(var))
lcl <- beta-qnorm(0.975)*SE 
ucl <- beta+qnorm(0.975)*SE
out<-cbind(exp(beta), exp(lcl), exp(ucl))
rownames(out)<-c("M","F")
out
# 0-0.5 vs 0
coef<-coef(cox_fit_7)[c(1,23)] %>% as.matrix(ncol=1)
vcov<-vcov(cox_fit_7)[c(1,23),c(1,23)]
var=t(c(1,1))%*%vcov%*%c(1,1)

beta <- c(coef(cox_fit_7)[1], sum(coef(cox_fit_7)[c(1,23)]))
SE <- c(coef(summary(cox_fit_7))[1,3], sqrt(var))
lcl <- beta-qnorm(0.975)*SE 
ucl <- beta+qnorm(0.975)*SE
out<-cbind(exp(beta), exp(lcl), exp(ucl))
rownames(out)<-c("M","F")
out
```

## subgroup analysis
```{r echo=FALSE, fig.height=4}
cox_fit_m<-coxph(Surv(AGEYRS, AGE_LAST, cancer_death)~BOOZE_q + BOOZE_q + as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN, data=data[data$SEX==1,])
summary(cox_fit_m)

cox_fit_f<-coxph(Surv(AGEYRS, AGE_LAST, cancer_death)~BOOZE_q + BOOZE_q + as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN, data=data[data$SEX==2,])
summary(cox_fit_f)

z <- (coef(cox_fit_m)[3] -coef(cox_fit_f)[3])/sqrt(vcov(cox_fit_m)[3,3] + vcov(cox_fit_f)[3,3])
p <- 1-pnorm(abs(z))
p
```
\newpage

# continuous BOOZE

## ~ BOOZE
```{r echo=FALSE, fig.height=4}
cox_fit_c1<-coxph(Surv(AGEYRS, AGE_LAST, cancer_death)~BOOZE, data=data)
summary(cox_fit_c1)
resic1<-cox.zph(cox_fit_c1)
resic1
plot(resic1, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_c1)[1])
```





\newpage

## ~ BOOZE + SEX + as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN
```{r echo=FALSE, fig.height=4}
cox_fit_c5<-coxph(Surv(AGEYRS, AGE_LAST, cancer_death)~BOOZE+SEX+ as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN, data=data)
summary(cox_fit_c5)
resic5<-cox.zph(cox_fit_c5)
resic5
plot(resic5, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_c5)[1])
```

\newpage

## ~ BOOZE + SEX + BOOZE_q * SEX + as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN
```{r echo=FALSE, fig.height=4}
cox_fit_c6<-coxph(Surv(AGEYRS, AGE_LAST, cancer_death)~BOOZE+SEX+BOOZE*SEX+ as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN, data=data)
summary(cox_fit_c6)
resic6<-cox.zph(cox_fit_c6)
resic6
plot(resic6, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_c6)[1])

coef<-coef(cox_fit_c6)[c(1,22)] %>% as.matrix(ncol=1)
vcov<-vcov(cox_fit_c6)[c(1,22),c(1,22)]
var=t(c(1,1))%*%vcov%*%c(1,1)

beta <- c(coef(cox_fit_c6)[1], sum(coef(cox_fit_c6)[c(1,22)]))
SE <- c(coef(summary(cox_fit_c6))[1,3], sqrt(var))
lcl <- beta-qnorm(0.975)*SE 
ucl <- beta+qnorm(0.975)*SE
out=cbind(exp(beta), exp(lcl), exp(ucl))
rownames(out)=c("M","F")
out
```

\newpage

## ~ BOOZE +BOOZE*strata(SEX)+ as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN, stratify by SEX 
```{r echo=FALSE, fig.height=4}
cox_fit_c7<-coxph(Surv(AGEYRS, AGE_LAST, cancer_death)~BOOZE +BOOZE*strata(SEX)+ strata(SEX) + as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN, data=data)
summary(cox_fit_c7)
resic7<-cox.zph(cox_fit_c7)
resic7
plot(resic7, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_c7)[1])

coef<-coef(cox_fit_c7)[c(1,21)] %>% as.matrix(ncol=1)
vcov<-vcov(cox_fit_c7)[c(1,21),c(1,21)]
var=t(c(1,1))%*%vcov%*%c(1,1)

beta <- c(coef(cox_fit_c7)[1], sum(coef(cox_fit_c7)[c(1,21)]))
SE <- c(coef(summary(cox_fit_c7))[1,3], sqrt(var))
lcl <- beta-qnorm(0.975)*SE 
ucl <- beta+qnorm(0.975)*SE
out=cbind(exp(beta), exp(lcl), exp(ucl))
rownames(out)=c("M","F")
out
```

## subgroup analysis
```{r}
cox_fit_m<-coxph(Surv(AGEYRS, AGE_LAST, cancer_death)~BOOZE  + as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN, data=data[data$SEX==1,])
summary(cox_fit_m)

cox_fit_f<-coxph(Surv(AGEYRS, AGE_LAST, cancer_death)~BOOZE + as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN, data=data[data$SEX==2,])
summary(cox_fit_f)

z <- (coef(cox_fit_m)[1] -coef(cox_fit_f)[1])/sqrt(vcov(cox_fit_m)[1,1] + vcov(cox_fit_f)[1,1])
p <- 1-pnorm(abs(z))
p
```

# ordinal BOOZE_q_ordinal

## ~ BOOZE_q_ordinal
```{r echo=FALSE, fig.height=4}
cox_fit_o1<-coxph(Surv(AGEYRS, AGE_LAST, cancer_death)~as.numeric(BOOZE_q), data=data)
summary(cox_fit_o1)
resio1<-cox.zph(cox_fit_o1)
resio1
plot(resio1, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_o1)[1])
```





\newpage

## ~ BOOZE_q_ordinal + SEX + as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN
```{r echo=FALSE, fig.height=4}
cox_fit_o5<-coxph(Surv(AGEYRS, AGE_LAST, cancer_death)~as.numeric(BOOZE_q)+SEX+ as.factor(RACE) + GRADES + as.factor(MARRY) + as.factor(SIZE) + AVGSMK + as.factor(SMSA) + URBAN, data=data)
summary(cox_fit_o5)
resio5<-cox.zph(cox_fit_o5)
resio5
plot(resio5, resid=FALSE, var=1, col=c("red","blue"))
abline(h=coef(cox_fit_o5)[1])
```

